---
title: "Assignment I - CompStat2023"
author: "Name of the Team"
date: "Team Member A, Team Member B, Team Member C"
output:
  pdf_document:
    toc: yes
    number_sections: true
  html_document:
    theme: readable
    toc: yes
    code_folding: show
    df_print: paged
---


```{r message=FALSE, warning=FALSE, include=FALSE}
library(magrittr)
library(dplyr)
library(ggplot2)
library(gridExtra)
library(ggpubr)
library(bookdown)
library(plotly)
library(rstanarm)
library(Cairo)
library(rgl)
library(plot3D)
library(effects)
library(MASS)
library(car)
library(faraway)
library(leaps)
library(ROCR)
library(RColorBrewer)
library(ISLR2)
library(e1071)
library(class)
library(reshape2)
library(VGAM)
library(sads)
```


# Montecarlo Simulation II 
## Point a)

\text{To prove that the random variable } $Y=U^{-\frac{1}{\gamma}}$ \text{, where U is Uniform in $(0, 1)$, follows a Pareto distribution, we need to show that its probability density function (PDF) matches the PDF of a Pareto distribution.}

\text{Let's begin by finding the cumulative distribution function (CDF) of } Y. \text{ We have:}\\

$$
F_Y(y) = P(Y \leq y) = P\left(U^{-1/\gamma} \leq y\right)\\
$$
\text{Taking the reciprocal of both sides, we get:}\\
$$
P\left(\frac{1}{U^{1/\gamma}} \leq \frac{1}{y}\right)\\
$$
\text{Since } U \text{ follows a standard uniform distribution on the interval } (0, 1), \text{ the above expression is equivalent to:}\\
$$
P\left(U^{1/\gamma} \geq y^{-1}\right) = P\left(U \geq y^{-\gamma}\right)\\
$$
\text{Since } U \text{ is uniformly distributed, the probability } $P(U \geq y^{-\gamma})$ \text{ is equal to } $1 - P(U < y^{-\gamma})$. \text{ Therefore, the CDF of } Y \text{ can be written as:}\\
$$
F_Y(y) = 1 - P(U < y^{-\gamma})\\
$$
\text{The probability } $P(U < y^{-\gamma})$ \text{ represents the CDF of a standard uniform distribution evaluated at } $y^{-\gamma}$. \text{ Since the CDF of a standard uniform distribution is } $F_U(u) = u$ \text{ for } $0 < u < 1$, \text{ we have:}\\
$$
F_Y(y) = 1 - y^{-\gamma}\\
$$
\text{Now, let's differentiate the CDF to obtain the PDF of } Y:\\
$$
f_Y(y) = \frac{d}{dy} F_Y(y) = \gamma y^{-\gamma-1}\\
$$
\text{Comparing this PDF with the PDF of a Pareto distribution, we can see that they are identical:}\\
$$
f_Y(y) = \gamma y^{-\gamma-1} = \gamma x^{-(\gamma+1)}\\
$$
\text{Therefore, we have shown that the random variable } $Y=U^{-\frac{1}{\gamma}}$ \text{ follows a Pareto distribution with parameter } $\gamma$.\\

## point b)

\text{To find the distribution of } $Y = \log(X)$ \text{, where } X \text{ follows a Pareto distribution with parameter } $\gamma$, \text{ we need to determine the probability density function (PDF) of } Y.\\

\text{Let's start by expressing the relationship between } X \text{ and } Y \text{ using the transformation function:}\\
$$
Y = \log(X)\\
$$
\text{To find the PDF of } Y, \text{ we can use the cumulative distribution function (CDF) method. Let's begin by finding the CDF of } Y.\\

\text{We have:}\\
$$
F_Y(y) = P(Y \leq y) = P(\log(X) \leq y)\\
$$
\text{Taking the exponential of both sides, we get:}\\
$$
P(X \leq e^y)\\
$$
\text{Since } X \text{ follows a Pareto distribution, its CDF is given by:}\\
$$
F_X(x) = 1 - \left(\frac{1}{x}\right)^\gamma \text{ for } x \geq 1\\
$$
\text{Using the CDF of } X, \text{ we can express the CDF of } Y \text{ as follows:}\\
$$
F_Y(y) = P(X \leq e^y) = 1 - \left(\frac{1}{e^y}\right)^\gamma = 1 - e^{-\gamma y}\\
$$
\text{To find the PDF of } Y, \text{ we differentiate the CDF with respect to } y:\\
$$
f_Y(y) = \frac{d}{dy} F_Y(y) = \gamma e^{-\gamma y}\\
$$
\text{Therefore, the PDF of } Y \text{ is given by:}\\
$$
f_Y(y) = \gamma e^{-\gamma y}\\
$$
\text{In conclusion, when } $Y = \log(X)$ \text{ and } X \text{ follows a Pareto distribution with parameter } $\gamma$, \text{ the resulting distribution of } Y \text{ is an exponential distribution with parameter } $\gamma$.\\


## point c)

```{r}
set.seed(123)

# Function to generate samples from a Pareto distribution
pareto_sampler <- function(n, gamma) {
  u <- runif(n)  # Generate n samples from a uniform distribution
  x <- (1/u)^(1/gamma)  # Transform the uniform samples to follow a Pareto distribution
  return(x)
}

# Function to generate samples from the logarithmic distribution
logarithmic_sampler <- function(n, gamma) {
  x <- pareto_sampler(n, gamma)  # Generate n samples from a Pareto distribution
  y <- log(x)  # Transform the Pareto samples to follow a logarithmic distribution
  return(y)
}

# Parameters
n <- 1000  # Number of samples
gamma <- 2  # Parameter of the Pareto distribution

# Generate samples from the Pareto and logarithmic distributions
x_samples <- pareto_sampler(n, gamma)
y_samples <- logarithmic_sampler(n, gamma)

# Plot histograms
par(mfrow = c(2, 1))
hist(x_samples, breaks = 30, main = "Pareto Distribution", xlab = "X", freq = FALSE, xlim = c(0, max(x_samples)))
curve(dpareto(x, scale=1, shape = 4))
hist(y_samples, breaks = 30, main = "Exponential Distribution", xlab = "Y", freq = FALSE, xlim = c(min(y_samples), max(y_samples)))
curve(dexp(x, rate=4))

# Plot density plots
curve((gamma * x^-(gamma + 1)), from = 0, to = max(x_samples), col = "blue", lwd = 2, ylab = "Density", main = "Pareto Distribution")
curve(dexp(x, gamma), from = min(y_samples), to = max(y_samples), col = "red", lwd = 2, ylab = "Density", main = "Logarithmic Distribution")
```

## point d)
```{r}
# Analytical approach
gamma <- 2  # Pareto parameter
p_analytical <- 1 - (1/5)^gamma

# Monte Carlo estimate
num_samples <- 1000000  # Number of samples
samples <- rpareto(num_samples,scale=1, shape=gamma)
p_mc <- mean(num_samples < 5)

# Output results
p_analytical
p_mc
```
